---
title: "cm007 Exercises: Practice with `dplyr`"
output: 
  html_document:
    keep_md: true
    theme: paper
---

<!---The following chunk allows errors when knitting--->

```{r allow errors, echo = FALSE}
knitr::opts_chunk$set(error = TRUE, warning = FALSE)

# and, also install the package
# install.packages("tsibble")

```

```{r}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(gapminder))
suppressPackageStartupMessages(library(tsibble))
```


This worksheet contains exercises aimed for practice with `dplyr`. 


1. (a) What's the minimum life expectancy for each continent and each year? (b) Add the corresponding country to the tibble, too. (c) Arrange by min life expectancy.

```{r}
temp <- gapminder %>% 
  group_by(continent, year) %>% 
  summarize(min_life = min(lifeExp),
            Country = country[lifeExp==min_life]) %>%
  arrange(min_life)
DT::datatable(temp)
```

2. Calculate the growth in population since the first year on record _for each country_ by rearranging the following lines, and filling in the `FILL_THIS_IN`. Here's another convenience function for you: `dplyr::first()`. 



```{r}
gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(rel_growth = pop - first(pop)) %>%
  DT::datatable()
```




3. Determine the country that experienced the sharpest 5-year drop (which is from one-data point) in life expectancy, in each continent, sorted by the drop, by rearranging the following lines of code. Ensure there are no `NA`'s. Instead of using `lag()`, use the convenience function provided by the `tsibble` package, `tsibble::difference()`:

_difference is same as lifeExp-lag(lifeExp)_

```{r}
gapminder %>% 
  group_by(country) %>% 
  arrange(year) %>% 
  mutate(inc_life_exp = tsibble::difference(lifeExp)) %>% 
  drop_na() %>% 
  ungroup() %>% 
  group_by(continent) %>% 
  filter(inc_life_exp == min(inc_life_exp)) %>% 
  arrange(inc_life_exp) %>% 
  knitr::kable()

```


